---
title: "LiteartureReviewScript_Adhikari2022"
author: "Biraj Adhikari"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

requirements
```{r}
library(networkD3) #for sankey Diagrams
library(dplyr)
library(ggplot2)
library(tidyr) #for the separate_rows function
library(ggpubr) #to combine multiple ggplot objects
```


# Figure 3: Temporal Trends in Research and Methods
```{r reading data}
data = read.csv('ForFig3_ResearchTrends_Methods.csv', header = TRUE, sep = ",")
```  

Making adjustments for the figures
```{r fig 3 data adjustments}
researchtrends <- data %>%
  group_by(Year) %>% summarize(n = n()) 
researchtrends$cumsum <- cumsum(researchtrends$n)

pies <- data %>% mutate(time.period = case_when(
  Year < 2008 ~ "First",
  Year >= 2008 & Year < 2015 ~ "Second",
  Year >=2015 & Year < 2018 ~ "Third",
  TRUE ~ "Fourth"
))
pies$time.period <- factor(pies$time.period, levels = c("First", "Second", "Third", "Fourth"))


pies <- pies %>% separate_rows(Method, sep = ", ") %>%
  group_by(time.period, Method) %>% summarize(Count = n())
pies$Method = factor(pies$Method, levels = c("Biophysical", "Monetary", "Socio-cultural"))

```

Creating a temporal trends in research figure
```{r fig 3(a)}
researchtrends_fig = ggplot(data = researchtrends, aes(x = Year, y = n)) + 
  
  #bar chart for publications per year
  geom_bar(color = "#608000", fill = "#608000", stat = "identity") + 
  
  #line/point plots for cumulative publications
  geom_line(aes(x = Year, y = cumsum/5, group = 1), size = 1,  color = "#ff9900") + 
  geom_point(aes(x = Year, y = cumsum/5), color = "#e68a00") + 
  
  #vertical lines indicating divides using natural Jenks
  geom_vline(xintercept = c(1994.5, 2007.5, 2014.5, 2017.5, 2020.5), 
             linetype = 2) + 
  
  #adjustments in axes
  labs( x = "Year") + 
  scale_x_continuous(breaks = seq(1995, 2020, 2), limits = c(1994.5, 2020.5)) + 
  scale_y_continuous("Number of publications", sec.axis = sec_axis(~. * 5, name = "Cumulative Publications")) + 
 
  #adjustment in theme
   theme(axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1), 
         legend.position = "bottom",
         panel.background = element_rect(fill = "white"),
         panel.grid.major = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"),
         panel.grid.minor = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"))

researchtrends_fig
```  

Creating the pie charts for methods
```{r fig 3(b)}
#1st Pie chart
pie1_fig = pies %>% filter(time.period == "First") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 
    #geom_text(aes(label = paste(round(Count / sum(Count) * 100, 1), "%")),
           # position = position_stack(vjust = 0.5), size = 2.5) +
  labs(title = "1995-2007") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie1_fig


#2nd Pie Chart
pie2_fig = pies %>% filter(time.period == "Second") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 
    #geom_text(aes(label = paste(round(Count / sum(Count) * 100, 1), "%")),
           # position = position_stack(vjust = 0.5), size = 2.5) +
  labs(title = "2007-2014") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie2_fig

#3rd Pie Chart
pie3_fig = pies %>% filter(time.period == "Third") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 
    #geom_text(aes(label = paste(round(Count / sum(Count) * 100, 1), "%")),
           # position = position_stack(vjust = 0.5), size = 2.5) +
  labs(title = "2015-2017") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie3_fig

#4th Pie Chart
pie4_fig = pies %>% filter(time.period == "Fourth") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 
    #geom_text(aes(label = paste(round(Count / sum(Count) * 100, 1), "%")),
           # position = position_stack(vjust = 0.5), size = 2.5) +
  labs(title = "2018-2020") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie4_fig
```  
Combining figures using facet grid

```{r combining 3(a) and 3(b), saving the plot}
png(filename = "Figure3.png",            
  width = 2400,        
  height = 2100,
  res = 300,  
  
  ) 

ggarrange(ggarrange(pie1_fig, pie2_fig, pie3_fig, pie4_fig, nrow = 1, ncol = 4, widths = c(1.3,0.8,0.5,0.5),  common.legend = TRUE, legend = "top"), researchtrends_fig, nrow = 2, ncol =1, heights = c(1, 3))

dev.off()
```  

```{r removing variables}
rm(list = ls())
```  

# Fig 5: Trends in NCP  
Figure 5 draws data from 2 files:  
1. ForFig5_NCPCount.csv: For the occurrence of NCP in the 140 articles  
2/ ForFig5_6_DriversTrends.csv: For the occurrences of positive and negative trends in NCP in the 140 articles   
```{r data add and adjustments}
#the ncpcount variable is used to make the overall count linegraph in Fig5
ncpcount <- read.csv("ForFig5_NCPCount.csv", header = TRUE, sep = ",")

#making this variable compatible with ggplot
ncpcount <- ncpcount[nrow(ncpcount),-1]
ncpcount<- data.frame(t(ncpcount))
ncpcount<- ncpcount %>% mutate(NCP = rownames(ncpcount))
rownames(ncpcount) <- c(1:nrow(ncpcount))
colnames(ncpcount) <- c("Count", "NCP")

#adding the labels
ncpcount <- ncpcount %>% 
mutate(Label = c("Habitat creaiton and maintenance (HAB)",
                 "Pollination and dispersal of seeds and other propagules (POL)",
                 "Regulation of air quality (AIR)",
                 "Regulation of climate (CLI)",
                 "Regulation of freshwater quantity, location and timing (WQN)",
                 "Regulation of freshwater and coastal water quality (WQL)",
                 "Formation, protection and decontamination of soils and sediments (SOI)",
                 "Regulation of hazards and extreme events (HAZ)",
                 "Regulation of detrimental organisms and biological processes (ORG)",
                 "Energy (NRG)",
                 "Food and feed (FOD)",
                 "Materials, companionship and labor (MAT)",
                 "Medicinal, biochemical and genetic resources (MED)",
                 "Learning and inspiration (INS)",
                 "Physical and psychological experiences (EXP)",
                 "Supporting identities (IDE)",
                 "Maintenance of options (OPT)"
                 
))

#this grp column is necessary for ggplot2
ncpcount <- ncpcount %>% mutate(grp = rep(1, nrow(ncpcount)))

#the labels of ncpcount is useful for another variable, so we store it for now
for_label <- ncpcount %>% select(NCP, Label)

#the ncp trend variable is used to create stacked bar charts of counts of NCP trends
ncptrend <-read.csv("ForFig5_6_DriversTrends.csv")
ncptrend <- ncptrend %>% 
  filter (NCP != "NA") %>% 
  group_by(NCP, Trend) %>%
  summarize(Count = round(sum(TrendCount)))

#we use the label of ncpcount for ncptrend too
ncptrend <- merge(ncptrend, for_label, by = "NCP")
rm(for_label)

#creating percentages to positive and negative trends
ncptrend<- ncptrend %>% group_by(Label, Trend) %>% summarize(TotalCount = sum(Count))
ncptrend <- ncptrend %>% mutate(Percent = round(TotalCount/sum(TotalCount), 2))

#factorizing the Trend column
ncptrend$Trend = factor(ncptrend$Trend, levels = c("Positive", "Negative"))
```  

Making the figure::
```{r ncp trends figure}
figure = ggplot() +
  #line graph for total number of occurrences for each NCP
  geom_line(data = ncpcount, 
            aes(x = reorder(Label, Count), y = Count, group = grp, color = '#003366'), 
            size = 1) + 
  #adding points within this line
  geom_point(data = ncpcount, 
              aes(x = Label, y = Count), color = '#003366', size = 1.2) + 
  #assigning colors again so that it shows up in the legend
  scale_colour_manual(name = "", values =c('#003366'), labels = c('Total Count')) + 
 
  
  #adding stacked bar chart for trends in positive and negative trends
  geom_col(data = ncptrend, 
           aes(x = Label, y = TotalCount, fill = Trend), width = 0.6, position = "stack") +
  
  #adding percentage values for negative trends for each NCP
  geom_text(data = ncptrend[ncptrend$Trend == "Negative",], 
          aes(x = Label, y = TotalCount, label=paste0(Percent*100,"%"), group = Trend),
          position=position_dodge(width=0.6), vjust= 1,  inherit.aes = TRUE, size = 2.5) + 
  
  #customizing colors for staked chart
  scale_fill_manual(values = c("#a6d854", "#fc8d62")) + 
  

 
   theme(
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.6, 'cm'),
    legend.text = element_text(size = 7.5),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text = element_text(size = 7.5),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"),
    panel.grid.minor = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0")
  ) + 
  
  scale_y_continuous(limits = c(0,85)) + 
  
  xlab("NCPs") +
  ylab("Count")+
  coord_flip() 

#saving to file
png(filename = "Figure5.png",            
  width = 2300,        
  height = 1500,
  res = 300,  
  ) 

figure
```  

```{r removing variables for figure 5}
rm(list = ls())
```


# Figure 6: Drivers of Change  
Note: We combine positive and negative drivers of change manually
## Positive Drivers of Change  
```{r}

```

