---
title: "LiteartureReviewScript_Adhikari2022"
author: "Biraj Adhikari"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

requirements
```{r}
library(networkD3) #for sankey Diagrams
library(dplyr)
library(ggplot2)
library(tidyr) #for the separate_rows function
library(ggpubr) #to combine multiple ggplot objects
```


# Figure 3: Temporal Trends in Research and Methods
```{r reading data, figure 3}
data = read.csv('ForFig3_ResearchTrends_Methods.csv', header = TRUE, sep = ",")
```  

Making adjustments for the figures
```{r fig 3 data adjustments}
researchtrends <- data %>%
  group_by(Year) %>% summarize(n = n()) 
researchtrends$cumsum <- cumsum(researchtrends$n)

pies <- data %>% mutate(time.period = case_when(
  Year < 2008 ~ "First",
  Year >= 2008 & Year < 2015 ~ "Second",
  Year >=2015 & Year < 2018 ~ "Third",
  TRUE ~ "Fourth"
))
pies$time.period <- factor(pies$time.period, levels = c("First", "Second", "Third", "Fourth"))


pies <- pies %>% separate_rows(Method, sep = ", ") %>%
  group_by(time.period, Method) %>% summarize(Count = n())
pies$Method = factor(pies$Method, levels = c("Biophysical", "Monetary", "Socio-cultural"))

```

Creating a temporal trends in research figure
```{r fig 3(a)}
researchtrends_fig <- ggplot(data = researchtrends, aes(x = Year, y = n)) + 
  
  #bar chart for publications per year
  geom_bar(color = "#608000", fill = "#608000", stat = "identity") + 
  
  #line/point plots for cumulative publications
  geom_line(aes(x = Year, y = cumsum/5, group = 1), size = 1,  color = "#ff9900") + 
  geom_point(aes(x = Year, y = cumsum/5), color = "#e68a00") + 
  
  #vertical lines indicating divides using natural Jenks
  geom_vline(xintercept = c(1994.5, 2007.5, 2014.5, 2017.5, 2020.5), 
             linetype = 2) + 
  
  #adjustments in axes
  labs( x = "Year") + 
  scale_x_continuous(breaks = seq(1995, 2020, 2), limits = c(1994.5, 2020.5)) + 
  scale_y_continuous("Number of publications", sec.axis = sec_axis(~. * 5, name = "Cumulative Publications")) + 
 
  #adjustment in theme
   theme(axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1), 
         legend.position = "bottom",
         panel.background = element_rect(fill = "white"),
         panel.grid.major = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"),
         panel.grid.minor = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"))

researchtrends_fig
```  

Creating the pie charts for methods
```{r fig 3(b)}
#1st Pie chart
pie1_fig <- pies %>% filter(time.period == "First") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 
  labs(title = "1995-2007") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie1_fig


#2nd Pie Chart
pie2_fig <- pies %>% filter(time.period == "Second") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 

  labs(title = "2007-2014") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie2_fig

#3rd Pie Chart
pie3_fig <- pies %>% filter(time.period == "Third") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 

  labs(title = "2015-2017") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie3_fig

#4th Pie Chart
pie4_fig <- pies %>% filter(time.period == "Fourth") %>% 
ggplot(aes(x = "", y = Count, fill = Method)) +
  geom_bar(stat = "identity", width = 1) + 
  scale_fill_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) + 

  labs(title = "2018-2020") + 
  
  coord_polar("y", start = 0) + 
  
  theme_void() +
  theme(legend.position = "none", 
        
        plot.title = element_text(hjust = 0.5, vjust = -3.5, size = 10))

pie4_fig
```  
Combining figures using facet grid

```{r combining 3(a) and 3(b), saving the plot}
figure <- ggarrange(ggarrange(pie1_fig, pie2_fig, pie3_fig, pie4_fig, nrow = 1, ncol = 4, widths = c(1.3,0.8,0.5,0.5),  common.legend = TRUE, legend = "top"), researchtrends_fig, nrow = 2, ncol =1, heights = c(1, 3))

png(filename = "Figure3.png",            
  width = 2400,        
  height = 2100,
  res = 300  
  ) 

figure

dev.off()

figure
```  

```{r removing variables}
rm(list = ls())
```  

# Fig 5: Trends in NCP  
Figure 5 draws data from 2 files:  
1. ForFig5_NCPCount.csv: For the occurrence of NCP in the 140 articles  
2/ ForFig5_6_DriversTrends.csv: For the occurrences of positive and negative trends in NCP in the 140 articles   
```{r data add and adjustments}
#the ncpcount variable is used to make the overall count linegraph in Fig5
ncpcount <- read.csv("ForFig5_NCPCount.csv", header = TRUE, sep = ",")

#making this variable compatible with ggplot
ncpcount <- ncpcount[nrow(ncpcount),-1]
ncpcount<- data.frame(t(ncpcount))
ncpcount<- ncpcount %>% mutate(NCP = rownames(ncpcount))
rownames(ncpcount) <- c(1:nrow(ncpcount))
colnames(ncpcount) <- c("Count", "NCP")

#adding the labels
ncpcount <- ncpcount %>% 
mutate(Label = c("Habitat creaiton and maintenance (HAB)",
                 "Pollination and dispersal of seeds and other propagules (POL)",
                 "Regulation of air quality (AIR)",
                 "Regulation of climate (CLI)",
                 "Regulation of freshwater quantity, location and timing (WQN)",
                 "Regulation of freshwater and coastal water quality (WQL)",
                 "Formation, protection and decontamination of soils and sediments (SOI)",
                 "Regulation of hazards and extreme events (HAZ)",
                 "Regulation of detrimental organisms and biological processes (ORG)",
                 "Energy (NRG)",
                 "Food and feed (FOD)",
                 "Materials, companionship and labor (MAT)",
                 "Medicinal, biochemical and genetic resources (MED)",
                 "Learning and inspiration (INS)",
                 "Physical and psychological experiences (EXP)",
                 "Supporting identities (IDE)",
                 "Maintenance of options (OPT)"
                 
))

#this grp column is necessary for ggplot2
ncpcount <- ncpcount %>% mutate(grp = rep(1, nrow(ncpcount)))

#the labels of ncpcount is useful for another variable, so we store it for now
for_label <- ncpcount %>% select(NCP, Label)

#the ncp trend variable is used to create stacked bar charts of counts of NCP trends
ncptrend <-read.csv("ForFig5_6_DriversTrends.csv")
ncptrend <- ncptrend %>% 
  filter (NCP != "NA") %>% 
  group_by(NCP, Trend) %>%
  summarize(Count = round(sum(TrendCount)))

#we use the label of ncpcount for ncptrend too
ncptrend <- merge(ncptrend, for_label, by = "NCP")
rm(for_label)

#creating percentages to positive and negative trends
ncptrend<- ncptrend %>% group_by(Label, Trend) %>% summarize(TotalCount = sum(Count))
ncptrend <- ncptrend %>% mutate(Percent = round(TotalCount/sum(TotalCount), 2))

#factorizing the Trend column
ncptrend$Trend <- factor(ncptrend$Trend, levels = c("Positive", "Negative"))
```  

Making the figure::
```{r ncp trends figure}
figure <- ggplot() +
  #line graph for total number of occurrences for each NCP
  geom_line(data = ncpcount, 
            aes(x = reorder(Label, Count), y = Count, group = grp, color = '#003366'), 
            size = 1) + 
  #adding points within this line
  geom_point(data = ncpcount, 
              aes(x = Label, y = Count), color = '#003366', size = 1.2) + 
  #assigning colors again so that it shows up in the legend
  scale_colour_manual(name = "", values =c('#003366'), labels = c('Total Count')) + 
 
  
  #adding stacked bar chart for trends in positive and negative trends
  geom_col(data = ncptrend, 
           aes(x = Label, y = TotalCount, fill = Trend), width = 0.6, position = "stack") +
  
  #adding percentage values for negative trends for each NCP
  geom_text(data = ncptrend[ncptrend$Trend == "Negative",], 
          aes(x = Label, y = TotalCount, label=paste0(Percent*100,"%"), group = Trend),
          position=position_dodge(width=0.6), vjust= 1,  inherit.aes = TRUE, size = 2.5) + 
  
  #customizing colors for staked chart
  scale_fill_manual(values = c("#a6d854", "#fc8d62")) + 
  

 
   theme(
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.6, 'cm'),
    legend.text = element_text(size = 7.5),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text = element_text(size = 7.5),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0"),
    panel.grid.minor = element_line(size = 0.15, linetype = "solid", colour = "#F0F0F0")
  ) + 
  
  scale_y_continuous(limits = c(0,85)) + 
  
  xlab("NCPs") +
  ylab("Count")+
  coord_flip() 

#saving to file
png(filename = "Figure5.png",            
  width = 2300,        
  height = 1500,
  res = 300,  
  ) 

figure

dev.off()

figure
```  

```{r removing variables for figure 5}
rm(list = ls())
```


# Figure 6: Drivers of Change  
Note: We combine positive and negative drivers of change manually
```{r reading data, figure 6}
overall_data <- read.csv("ForFig5_6_DriversTrends.csv")

overall_data$Indirect_Driver <- factor(overall_data$Indirect_Driver, levels = c("Demographic and Sociocultural", "Economic and Technological", "Institutions and Governance", "Conflicts"))
```    

Making the data adjustible to networkD3 function

```{r data cleaning Figure 6}
#extracting data for Indirect Driver -> Direct Driver
countdata_1 <- overall_data %>% filter(Indirect_Driver != "NA") %>% filter(Direct_Driver != "Directly to NCP") %>% filter(Direct_Driver != "NA") %>% group_by(Indirect_Driver, Direct_Driver, Trend) %>%  summarize(count = n())
colnames(countdata_1) = c("source", "target","type","count")

#extracting Indirect Drivers -> NCP
countdata_2 <- overall_data %>% 
  filter(Direct_Driver == "Directly to NCP") %>% 
  group_by(Indirect_Driver, NCP, Trend) %>%  
  summarize(count = n())
colnames(countdata_2) <- c("source", "target","type","count")

#extracting Direct Drivers -> Nature
countdata_3 <- overall_data %>% filter(Direct_Driver != "Directly to NCP") %>% filter(Direct_Driver != "NA") %>% filter(Nature != "NA") %>% group_by(Direct_Driver, Nature, Trend) %>%  summarize(count = n())
colnames(countdata_3) <- c("source", "target","type","count")

#extracting Nature -> NCP
countdata_4 <- overall_data %>% filter(Nature != "Directly to NCP") %>% filter(Nature != "NA") %>% filter(NCP != "NA") %>% group_by(Nature, NCP, Trend) %>%  summarize(count = n())
colnames(countdata_4) <- c("source", "target","type","count")
```

## Positive Drivers of Change
```{r plotting positive drivers}
links <- rbind(countdata_1, countdata_2, countdata_3, countdata_4)


links$group <- as.factor(c(rep("a", 7), rep("b",5), rep("c", 7), "d", rep("a", 6), rep("b",6), rep("c", 12), rep("e", 5), rep("f",8), rep("g", 5), rep("h", 10), "i", rep("j",5), rep("k", 9), rep("l", 19), rep("m",11), rep("n", 4), rep("o",4)))
links = filter(links, type == "Positive")

nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)


links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1

nodes$group <- as.factor(c("a", "b", "c", "f", "h", "j", "k", "l", "m", "n", "o", rep("p", 7), rep("p",4)))


my_color <-  'd3.scaleOrdinal() .domain(["a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p"]) .range(["#decbe4", "#d9d9d9","#fddaec", "#fb6a4a", "#F9A778", "#b3cde3", "#e5d8bd", "#ccffe6", "#ffd9cc", "#ffb3b3", "#FAE1AA", "#b2df8a", "#66cccc", "#FBEF8B","#978E9D ", "#CCFAAA"])'


# Make the Network
drivers_positive <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "count", NodeID = "name", colourScale = my_color, NodeGroup = "group", LinkGroup = "group",
              sinksRight = TRUE, nodeWidth = 15, nodePadding = 12, fontSize = 18, height = 500, width = 1000)

drivers_positive

```  

## Negative Drivers of Change
```{r plotting negative drivers}
links <- rbind(countdata_1, countdata_2, countdata_3, countdata_4)

links$group <- as.factor(c(rep("a", 7), rep("b",5), rep("c", 7), "d", rep("a", 6), rep("b",6), rep("c", 12), rep("e", 5), rep("f",8), rep("g", 5), rep("h", 10), "i", rep("j",5), rep("k", 9), rep("l", 19), rep("m",11), rep("n", 4), rep("o",4)))
links <- filter(links, type == "Negative")

nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)


links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1


nodes$group <- as.factor(c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j" , "k", "l", "m", "n" , "o", rep("p",14)))

my_color =  'd3.scaleOrdinal() .domain(["a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p"]) .range(["#decbe4", "#d9d9d9","#fddaec", "#fb6a4a", "#F9A778", "#b3cde3", "#e5d8bd", "#ccffe6", "#ffd9cc", "#ffb3b3", "#FAE1AA", "#b2df8a", "#66cccc", "#FBEF8B","#978E9D ", "#CCFAAA"])'


# Make the Network
drivers_negative <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "count", NodeID = "name", colourScale = my_color, NodeGroup = "group", LinkGroup = "group",
              sinksRight = TRUE, nodeWidth = 15, nodePadding = 12, fontSize = 18, height = 500, width = 1000)

drivers_negative
```  

```{r deleting files used for fig 6}
rm(list = ls())
```


# Fig 7: NCP - SDG linkages  
```{r adding data, fig. 7}
data <- read.csv("ForFig7_NCPSDGLinkages.csv")


data$Ecosystem <- factor(data$Ecosystem, 
                        levels = c("Forest", "Freshwater", "Farmland", "Grassland",
                                   "Agroforest", "Others"))

#first, we need to separate out the 1.5, 11.5, and 13.1 targets:
data <- separate_rows(data, SDG, sep = ", ")
#now, only retaining the goals, and removing the dots.
data$SDG <- gsub("\\..*", "", data$SDG)
#finally, adding "SDG" infornt of these numbers::
data$SDG <- paste("SDG ", data$SDG, sep = "")

#extracting data for Ecosystem -> NCP
countdata_1 <- data %>% filter(NCP_short != "NA") %>% group_by(Ecosystem, NCP_short, Effect) %>%  summarize(count = n())
colnames(countdata_1) <- c("source", "target","type","count")

#extracting NCP -> SDG
countdata_2 <- data %>% 
  filter(!is.na(NCP_short)) %>%
  filter(SDG != "SDG NA") %>%#weird category formed due to the paste function
 # filter(!is.na(SDG)) %>%
  group_by(NCP_short, SDG, Effect) %>%
  summarize(count = n())
colnames(countdata_2) = c("source", "target","type","count")
```

```{r making the figure}
links <- rbind(countdata_1, countdata_2)

nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)

links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
  
nodes$group <- as.factor(c("a", "b", "c", "d", "e", "f", "reg", "reg", "nmat", "mat",
                          "reg", "reg", "nmat", "nmat", "mat", "mat", "mat",
                          "nmat", "reg", "reg", "reg", "reg", "reg", rep("sdg", 12)))

#assigning color values based on (1) Individual Ecosystems, (2) NCP category and (3) SDGs               
my_color <-  'd3.scaleOrdinal() .domain(["a", "b", "c", "d", "e", "f", "reg", "mat", "nmat", "sdg", "Negative", "Positive"]) .range(["#b2df8a", "#66cccc","#FAE1AA", "#FBEF8B","#f2f2f2", "#978E9D",  "#CCFAAA", "#ffd6cc", "#79bcd2", "#ffebe6", "#fbb4ae", "#ccebc5"])'

# Make the Network
ncp2sdg <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "count", NodeID = "name", colourScale = my_color, NodeGroup = "group", 
              LinkGroup = "type", sinksRight = TRUE, nodeWidth = 15, nodePadding = 12, fontSize = 14)

ncp2sdg
```

```{r removing data Figure 7}
rm(list = ls())
```



